### 2023-06-30 scRNAseq/processing.Rmd

---
title: "data processing"
output: 
  html_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.lazy=FALSE, message=FALSE, warning=FALSE)
library(sctransform)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(cowplot)
library(ggrepel)
library(dplyr)
library(scRepertoire)
library(tidyr)
library(cluster)
library(plyr)
library(ComplexHeatmap)
library(dendextend)
options(future.globals.maxSize = 10 * 1024 * 1024^2)
path <- getwd()
```

we follow the [Seurat vignette](https://satijalab.org/seurat/v4.0/reference_mapping.html) and use their PBMC data as reference

```{r get_reference, fig.width=6,fig.height=5}
reference <- LoadH5Seurat(file.path(path, 'seurat', "pbmc_multimodal.h5seurat"))
```

load PBMC data; take all cells with less than 10% mitochondrial RNA and between 250 and 5000 genes

- K456985_PBMC
- K783496_PBMC
- K693104_PBMC
- K955601_PBMC
- P1_1_PBMC 
- P1_2_PBMC
- P3_1_PBMC
- P3_2_PBMC
- P8_1_PBMC
- P8_2_PBMC 
- P22_1_PBMC
- P22_2_PBMC
- P23_1_PBMC
- P23_2_PBMC
- P25_1_PBMC
- P25_2_PBMC
- P27_1_PBMC
- P27_2_PBMC
- P28_1_PBMC
- P28_2_PBMC
- P22_P23_K783496_CD4
- P8_P3_K456985_CD4
- P27_P28_K693104_CD4
- P12_1_PBMC
- P12_2_PBMC
- P26_1_PBMC
- P26_2_PBMC
- K496398_PBMC
- P12_P26_K496398_CD4

```{r get_data}
samples <- c('K456985_PBMC','K783496_PBMC','P1_1_PBMC','P22_1_PBMC','P22_P23_K783496_CD4','P23_2_PBMC','P25_2_PBMC','P27_2_PBMC',
             'P28_1_PBMC','P3_1_PBMC','P8_1_PBMC','P8_P3_K456985_CD4','K693104_PBMC','K955601_PBMC','P1_2_PBMC','P22_2_PBMC',
             'P23_1_PBMC','P25_1_PBMC','P27_1_PBMC','P27_P28_K693104_CD4','P28_2_PBMC','P3_2_PBMC','P8_2_PBMC',
             'P12_1_PBMC','P12_2_PBMC','P26_1_PBMC','P26_2_PBMC','K496398_PBMC','P12_P26_K496398_CD4')

pbmc <- list()
for (sample in samples) {
  dge <- Read10X_h5(file.path(path,'cellranger',paste0(sample,'_filtered_feature_bc_matrix.h5')))
  sobj <- CreateSeuratObject(counts = dge[['Gene Expression']], project=sample)
  sobj[['pct.mito']] <- PercentageFeatureSet(sobj, pattern='^MT-')
  sobj <- subset(sobj, (pct.mito < 10) & (nFeature_RNA <= 5000) & (nFeature_RNA >= 250))
  capture.output(sobj <- SCTransform(sobj, verbose = FALSE),file='/dev/null')
  cells <- colnames(sobj)
  row.names(dge[['Antibody Capture']]) <- gsub('_TotalSeqC','',row.names(dge[['Antibody Capture']]))
  hashtags <- grepl('^Hashtag_[1234]',row.names(dge[['Antibody Capture']]))
  sobj[["orig_ADT"]] <- CreateAssayObject(counts = dge[['Antibody Capture']][!hashtags,cells,drop=FALSE])
  sobj <- NormalizeData(sobj, normalization.method='CLR', assay='orig_ADT', verbose=FALSE, margin=2)
  if (grepl('CD4',sample)) {
    hts <- c("Hashtag_1","Hashtag_2","Hashtag_3","Hashtag_4")
    sobj[["HTO"]] <- CreateAssayObject(counts = dge[['Antibody Capture']][hts,cells,drop=FALSE])
    sobj <- NormalizeData(sobj, normalization.method='CLR', assay='HTO', verbose=FALSE)
    sobj <- HTODemux(sobj, assay='HTO', verbose=FALSE)
  }
  pbmc[[sample]] <- sobj
}
```

get some QC metrics

```{r QC,fig.width=10,fig.height=5}
md <- do.call(rbind, lapply(names(pbmc), function(x) pbmc[[x]]@meta.data[,c('nCount_RNA','nFeature_RNA','pct.mito')] %>%
                              dplyr::mutate(orig.ident=x))) %>%
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(orig.ident=paste0(orig.ident,' (n=',n(),')')) %>%
  gather(metric,value,-orig.ident)

ggplot(md,aes(x=orig.ident,y=value,fill=orig.ident)) + 
  #geom_boxplot(outlier.shape=NA) + 
  geom_violin(scale='width') +
  facet_wrap(~metric,scales='free_y') + 
  theme_classic() + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') + 
  labs(x='',y='')

```

project our cells onto this reference embedding using RNA data; predict cell type labels for the level 1, level 2 and level 3 clustering of the reference, and also antibody data ("what ADT data would be expect if our samples are like the reference?")

```{r project}
for (sample in samples) {
  anchors <- FindTransferAnchors(
    reference = reference,
    query = pbmc[[sample]],
    normalization.method = "SCT",
    reference.reduction = "spca",
    dims = 1:50,
    verbose=FALSE)
  pbmc[[sample]] <- MapQuery(
    anchorset = anchors,
    query = pbmc[[sample]],
    reference = reference,
    refdata = list(
      celltype.l1 = "celltype.l1",
      celltype.l2 = "celltype.l2",
      celltype.l3 = "celltype.l3",
      predicted_ADT = "ADT"
    ),
    reference.reduction = "spca", 
    reduction.model = "wnn.umap",
    verbose=FALSE)
}
```

identify and remove putative doublets (shown in pink), and cells with prediction score < .75

```{r doublets,fig.width=10,fig.height=12}
source(file.path(path,'my_doubletFinder.R'))
pbmc <- lapply(pbmc, function(sobj) {
  DefaultAssay(sobj) <- 'RNA'
  Idents(sobj) <- 'predicted.celltype.l1'
  homotypic.prop <- modelHomotypic(Idents(sobj))
  nExp_poi <- round(0.05*nrow(sobj@meta.data)) 
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  tmp <- my_doubletFinder(sobj, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE)
  tmp <- my_doubletFinder(tmp, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = "pANN")
  AddMetaData(sobj, tmp@meta.data[,'DF.classifications',drop=FALSE])
})

do.call(rbind,lapply(pbmc, function(sobj) table(sobj$DF.classifications))) %>%
  DT::datatable()

plots <- lapply(names(pbmc), function(sample)
  DimPlot(pbmc[[sample]], group.by='DF.classifications',
          cols=c('Singlet'='gray',
                 'Doublet'='deeppink')) +
    theme_void() + 
    labs(title=sample) + 
    theme(legend.position='none'))

plot_grid(plotlist=plots,ncol=5,align='vh')

pbmc.sub <- lapply(pbmc, function(x) subset(x, subset=(DF.classifications=='Singlet') & (predicted.celltype.l1.score >= .75)))
```

```{r save_samples}
for (sample in samples) {
  saveRDS(pbmc.sub[[sample]], file.path(path,'seurat',paste0(sample,'.rds')))
}
```

aggregate and demultiplex hashtags

```{r aggregate}
pbmc <- list()
for (lib in grep('PBMC',samples,value=TRUE)) {
  pbmc[[lib]] <- readRDS(file.path(path,'seurat',paste0(lib,'.rds')))
  pbmc[[lib]]$sample <- lib
}

study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))

assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Library.Name=gsub('p','P',Library.Name))

for (lib in grep('CD4',samples,value=TRUE)) {
  sobj <- readRDS(file.path(path,'seurat',paste0(lib,'.rds')))
  vireo <- read.table(file.path(path,'vireo',paste0(lib,'_donor_ids.tsv')),
                      header=1,row.names=1,as.is=TRUE)
  colnames(vireo) <- paste0('vireo_',colnames(vireo))
  sobj <- AddMetaData(sobj, vireo)  
  sobj$vireo_donor_id[is.na(sobj$vireo_donor_id)] <- 'none'
  samples <- assay %>% 
    dplyr::filter(Library.Name==lib) %>% 
    dplyr::select(Parameter.Value.Multiplex.sample.tag.,Sample.Name) %>%
    dplyr::mutate(ht=Parameter.Value.Multiplex.sample.tag.) %>%
    dplyr::mutate(ht=ifelse(ht=='','Negative',gsub('_','-',ht))) 
  samples <- setNames(as.character(samples$Sample.Name), samples$ht)
  sobj$sample <- revalue(sobj$hash.ID, samples)
  print(table(sobj$sample,sobj$vireo_donor_id))
  sobjs <- SplitObject(sobj,split.by='sample')
  for (sample in samples) {
    pbmc[[sample]] <- sobjs[[sample]]
  }
}
```

add clonotype information

add clonotype information (TCR and BCR where available; now with `filterMulti=TRUE`)

```{r get_clonotypes}
cols <- c('sample','orig.ident')
samples <- assay %>% 
  dplyr::filter(Library.Name %in% libraries) %>%
  dplyr::select(Library.Name,Sample.Name)

bcs <- lapply(pbmc, function(x) gsub('.*\\.','',Cells(x)))

tcr_list <- list()
bcr_list <- list()
for (lib in libraries) {
  infile <- file.path(path,'cellranger',paste0(lib,'vdj_t_filtered_contig_annotations.csv'))
  if (file.exists(infile)) {
    tcr <- read.csv(infile,
                    stringsAsFactors=FALSE) %>%
      dplyr::mutate(is_cell=is_cell=='true',
                    high_confidence=high_confidence=='true',
                    full_length=full_length=='true',
                    productive=productive=='true')
    for (sample in samples %>% dplyr::filter(Library.Name==lib) %>% dplyr::pull(Sample.Name)) {
      tcr_list[[sample]] <- tcr %>% dplyr::filter(barcode %in% bcs[[sample]])
    }
  }
  infile <- file.path(path,'cellranger',paste0(lib,'vdj_b_filtered_contig_annotations.csv'))
  if (file.exists(infile)) {
    bcr <- read.csv(infile,
                    stringsAsFactors=FALSE) %>%
      dplyr::mutate(is_cell=is_cell=='true',
                    high_confidence=high_confidence=='true',
                    full_length=full_length=='true',
                    productive=productive=='true')
    for (sample in samples %>% dplyr::filter(Library.Name==lib) %>% dplyr::pull(Sample.Name)) {
      bcr_list[[sample]] <- bcr %>% dplyr::filter(barcode %in% bcs[[sample]])
    }
  }
}

tcr_combined <- sapply(which(sapply(tcr_list, function(x) !is.null(x))),
                       function(i) combineTCR(tcr_list[[i]],
                                              samples=c(names(tcr_list)[[i]]),
                                              ID=c(names(tcr_list)[[i]]),
                                              cells='T-AB',
                                              filterMulti=TRUE))
names(tcr_combined) <- names(tcr_list)[sapply(tcr_list, function(x) !is.null(x))]

bcr_combined <- sapply(which(sapply(bcr_list, function(x) !is.null(x))),
                       function(i) combineBCR(bcr_list[[i]],
                                              samples=c(names(bcr_list)[[i]]),
                                              ID=c(names(bcr_list)[[i]])))
names(bcr_combined) <- names(bcr_list)[sapply(bcr_list, function(x) !is.null(x))]
```

```{r save_all}
saveRDS(pbmc,file.path(path,'seurat','PBMC_combined.rds'))
saveRDS(tcr_combined,file.path(path,'seurat','TCR_combined.rds'))
saveRDS(bcr_combined,file.path(path,'seurat','BCR_combined.rds'))
```

```{r sessionInfo}
sessionInfo()
```