### 2024-03-11 scRNAseq/paper_figures.Rmd
update figures for revision

---
title: "paper figures"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.lazy=FALSE, message=FALSE, warning=FALSE, dev=c('png','pdf'))
library(sctransform)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(cowplot)
library(ggrepel)
library(dplyr)
library(scRepertoire)
library(tidyr)
library(cluster)
library(plyr)
library(pheatmap)
library(dendextend)
library(ggpubr)
library(ComplexHeatmap)
library(RColorBrewer)
library(Binarize)
library(limma)
options(future.globals.maxSize = 10 * 1024 * 1024^2)
path <- getwd()
```

# setup

load our own PBMC data after projecting on to the reference

- K456985_PBMC
- K783496_PBMC
- K693104_PBMC
- K955601_PBMC
- P1_1_PBMC
- P1_2_PBMC
- P3_1_PBMC
- P3_2_PBMC
- P8_1_PBMC
- P8_2_PBMC
- P22_1_PBMC
- P22_2_PBMC
- P23_1_PBMC
- P23_2_PBMC
- P25_1_PBMC
- P25_2_PBMC
- P27_1_PBMC
- P27_2_PBMC
- P28_1_PBMC
- P28_2_PBMC
- P22_P23_K783496_CD4
- P8_P3_K456985_CD4
- P27_P28_K693104_CD4
- P12_1_PBMC
- P12_2_PBMC
- P26_1_PBMC
- P26_2_PBMC
- K496398_PBMC
- P12_P26_K496398_CD4

```{r get_data}
pbmc <- readRDS(file.path(path,'seurat','PBMC_combined.rds'))
tcr_combined <- readRDS(file.path(path,'seurat','TCR_combined.rds'))
bcr_combined <- readRDS(file.path(path,'seurat','BCR_combined.rds'))
CD4.integrated <- readRDS(file.path(path,'seurat','CD4_reclustered.rds'))
CD4.integrated <- RenameIdents(CD4.integrated, '12'='10')
Idents(CD4.integrated) <- as.factor(as.numeric(Idents(CD4.integrated)))
```

```{r get_metadata}
study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Library.Name=gsub('p','P',Library.Name))

timepoint_colors <- c('ctrl'='gray','Vedo_0d'='#ff7700','Vedo_6wk'='#882e72')
condition_colors <- c('HC'='black','UC'='#1965b0')
responder_colors <- c('R'='#34a048','NR'='#db2F26')
```

```{r normalize, cache=FALSE}
for (nn in names(pbmc)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
  pbmc[[nn]] <- NormalizeData(pbmc[[nn]],normalization.method='LogNormalize',scale.factor=10000,verbose=FALSE)
}
```

show all data in the reference embedding; match cluster granularity to CyTOF data;

```{r Fig_S2A,fig.width=4.5,fig.height=3.5}
ct_map <- c('CD8 TCM'='Mmry CD8+ T cells',
            'CD8 TEM'='Mmry CD8+ T cells',
            'CD8 Naive'='Naive CD8+ T cells', 
            'CD4 TCM'='Mmry CD4+ T cells',
            'CD4 TEM'='Mmry CD4+ T cells',
            'CD4 Naive'='Naive CD4+ T cells',
            'NK'='NK cells',
            'NK Proliferating'='NK cells',
            'NK_CD56bright'='NK cells',
            'gdT'='gd T cells',
            'MAIT'='MAIT cells',
            'B intermediate'='IgD+ B cells',
            'B naive'='IgD+ B cells',
            'B memory'='IgD- B cells',
            'CD14 Mono'='cMonocytes',
            'CD16 Mono'='ncMonocytes',
            'CD4 CTL'='Cytotoxic T cells',
            'CD4 Proliferating'='Mmry CD4+ T cells',
            'CD8 Proliferating'='Mmry CD8+ T cells',
            'cDC1'='cDCs',
            'cDC2'='cDCs',
            'pDC'='pDCs',
            'Doublet'='other',
            'Eryth'='other',
            'dnT'='other',
            'ASDC'='other',
            'HSPC'='other',
            'ILC'='ILCs',
            'Plasmablast'='Plasmablast',
            'Platelet'='other',
            'Treg'='Tregs')

celltype_colors <- c('IgD+ B cells'='#E57A3B',
                     'IgD- B cells'='#A57496',
                     'Cytotoxic T cells'='darkolivegreen1',
                     'Naive CD4+ T cells'='#399649',
                     'Mmry CD4+ T cells'='#D3317D',
                     'Naive CD8+ T cells'='#AFD087',
                     'Mmry CD8+ T cells'='#D685AC',
                     'cDCs'='#E47970',
                     'pDCs'='#9B743B',
                     'cMonocytes'='#295D98',
                     'ncMonocytes'='#5999A2',
                     'NK cells'='#8FCBBA',
                     'other'='gray30',
                     'Plasmablast'='#D9A53C',
                     'gd T cells'='#7A3163',
                     'MAIT cells'='#F0AD69',
                     'Tregs'='darkolivegreen3',
                     'ILCs'='plum1')

pbmc_meta <- do.call(rbind,
                     lapply(pbmc[grepl('PBMC',names(pbmc))], 
                            function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                             'orig.ident','sample',
                                                             'predicted.celltype.l1',
                                                             'predicted.celltype.l2',
                                                             'predicted.celltype.l3')) %>%
                              tibble::rownames_to_column('cell') %>%
                              dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_meta, aes(x=refUMAP_1,y=refUMAP_2,color=celltype)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=celltype_colors) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

```{r Fig_S2B,fig.width=4.5,fig.height=3.5}
ggplot(pbmc_meta %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=Patient.ID)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

create heatmap of ADT expression for these celltypes

```{r Fig_S2C,fig.width=7,fig.height=4.5, cache=FALSE}
ABs <- c("CD3","CD8a","CD40","CD19","CD20","CD45RA","CD38","CD69","HLA-DR",
         "CD11c","CD123","CD16","CD161","CD1c","CD14","CD25","CD127","CD56","CD4","CD45")

for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'orig_ADT'
}

ADT <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c(ABs,'orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(pbmc_meta[,c('cell','celltype')],by='cell') %>%
  dplyr::mutate_if(is.numeric,  scales::rescale, to=c(0,1)) %>%
  dplyr::group_by(celltype) %>%
  dplyr::summarise_if(is.numeric, median)

pheatmap(ADT %>% tibble::column_to_rownames('celltype'),
         color=colorRampPalette(c("white","gray80","gray20","black"))(100),
         annotation_row = data.frame(celltype=as.factor(ADT$celltype),row.names=ADT$celltype),
         annotation_colors = list(celltype=celltype_colors[ADT$celltype]),
         annotation_legend=FALSE,
         annotation_names_row=FALSE,
         name='scaled median expression')
```

compare PBMC composition between the time points

```{r Fig_S2D, fig.width=7,fig.height=4}
#library(lme4)
library(gtools)

cols <- c('sample','orig.ident','predicted.celltype.l2')
all_frac <- do.call(rbind,lapply(pbmc, function(x) x[[cols]])) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map)) %>%
  dplyr::group_by(sample,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n)) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

pbmc_frac <- all_frac[all_frac$Material=='PBMC',]
pbmc_frac$Timepoint <- factor(pbmc_frac$Timepoint, levels=c('Vedo_0d','Vedo_6wk','ctrl'))

pl1 <- pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('ctrl','Vedo_0d')) %>%
  dplyr::mutate(condition=factor(ifelse(Diagnosis=='control','HC','UC'),levels=c('HC','UC'))) %>%
  ggplot(aes(x=condition,y=frac,fill=condition,color=condition)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('UC','HC')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=condition_colors,limits=force) +
  scale_color_manual(values=condition_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='')

pg <- ggplotGrob(pl1)
for(i in which(grepl("strip-b", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl1
```

```{r Fig_S3A,fig.width=4.5,fig.height=4, cache=FALSE}
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
}

RNA <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('ITGA4','ITGB7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','ctrl')) %>%
  gather(gene,expression,c(ITGA4,ITGB7)) %>%
  dplyr::arrange(gene,expression)

ggplot(RNA, aes(x=refUMAP_1,y=refUMAP_2,color=expression)) +
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  facet_grid(gene~Diagnosis,switch='y') +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       #lim=c(1.e-6,5), 
                       #breaks=c(1.e-6,5),
                       #labels=c('low','high'),
                       oob=scales::squish,
                       guide=guide_colorbar(title.position='right',
                                            title.theme=element_text(angle=90,hjust=.5,vjust=.5))) + 
  theme(legend.key.size=unit(3.5,'mm')) +
  labs(color='')
```

```{r Fig_S3D,fig.width=7,fig.height=3}
pl <- RNA %>% 
  spread(gene,expression) %>% 
  dplyr::group_by(orig.ident,Patient.ID, celltype,Timepoint,Diagnosis) %>% 
  dplyr::summarise(pos=mean((ITGA4 > 0) & (ITGB7 > 0))) %>%
  dplyr::filter(Timepoint %in% c("Vedo_0d","ctrl")) %>%
  dplyr::mutate(condition=factor(ifelse(Diagnosis=='control','HC','UC'),levels=c('HC','UC'))) %>%
  ggplot(aes(x=condition,y=pos,fill=condition,color=condition)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('UC','HC')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=condition_colors,limits=force) +
  scale_color_manual(values=condition_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='fraction ITGA4/ITGB7\ndouble-positive cells',fill='',color='')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip-b", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
```

differential expression of ITGA4 and ITGB7

```{r Fig_S3B,fig.width=5,fig.height=4}
RNA <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('ITGA4','ITGB7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  gather(gene,expression,c(ITGA4,ITGB7)) %>%
  dplyr::arrange(gene,expression)

ggplot(RNA %>%
         dplyr::group_by(gene,Timepoint,celltype) %>%
         dplyr::summarise(pct.expressed=100*mean(expression > 0),
                          mean_expression=mean(expression)), 
       aes(x=Timepoint,y=celltype,color=mean_expression,size=pct.expressed)) +
  geom_point() +
  facet_wrap(~gene,ncol=2) +
  theme_bw() +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       #lim=c(1.e-6,1.3),
                       #breaks=c(1.e-6,1.3),
                       #labels=c('low','high'),
                       oob=scales::squish) +
  theme(legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        axis.text.x=element_text(angle=90,hjust=1,vjust=0)) +
  labs(x='',y='',color='')
```

```{r Fig_S3C,fig.width=3.75,fig.height=4, cache=FALSE}
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'orig_ADT'
}

ADT <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('Integrin-b7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  gather(gene,expression,`Integrin-b7`) %>%
  dplyr::arrange(gene,expression)

threshold <- binarize.kMeans(ADT$expression)@threshold

ggplot(ADT %>%
         dplyr::group_by(gene,Timepoint,celltype) %>%
         dplyr::summarise(pct.expressed=100*mean(expression > threshold),
                          mean_expression=mean(expression)), 
       aes(x=Timepoint,y=celltype,color=mean_expression,size=pct.expressed)) +
  geom_point() +
  facet_wrap(~gene,ncol=2) +
  theme_bw() +
  scale_color_gradient(low='honeydew1',high='forestgreen', 
                       na.value='gray80', 
                       # lim=c(1.e-6,1.3),
                       # breaks=c(1.e-6,1.3),
                       # labels=c('low','high'),
                       oob=scales::squish) +
  theme(legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        axis.text.x=element_text(angle=90,hjust=1,vjust=0)) +
  labs(x='',y='',color='')
```

show where we detect clonotypes (not yet filtered for CD4/CD8 T cells!)

```{r Fig_S4B,fig.width=4.5,fig.height=3.5, cache=FALSE}
pbmc.T <- lapply(names(pbmc), function(sample) {
  bcs <- tcr_combined[[sample]]$barcode
  cells <- paste0(sample,'_',sample,'_',Cells(pbmc[[sample]]))
  tmp <- RenameCells(pbmc[[sample]], new.names=cells)
  combineExpression(tcr_combined[[sample]][bcs %in% cells,],tmp, cloneCall="aa")
  })
names(pbmc.T) <- names(pbmc)

pbmc_T_meta <- do.call(rbind,
                       lapply(pbmc.T,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
                                #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_T_meta %>% dplyr::mutate(TCR_detection=ifelse(is.na(CTaa),'no TCR','TCR')) %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=TCR_detection)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=c('TCR'='lightblue','no TCR'='grey'),limits=force) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))

cloneType_colors <- c('hyperexpanded'='plum4',
                      'large'='plum3',
                      'medium'='plum2',
                      'small'='plum1')

ggplot(pbmc_T_meta %>% dplyr::mutate(cloneType=factor(tolower(gsub('Single','Small',gsub(' .*','',cloneType))),
                                                      levels=c('small','medium','large','hyperexpanded'))) %>% dplyr::sample_frac(1L),
       aes(x=refUMAP_1,y=refUMAP_2,color=cloneType)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=cloneType_colors,na.value='gray80') +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

<!-- ```{r clonotype_celltypes_TCR,fig.width=6.5,fig.height=10} -->
<!-- tmp <- list() -->
<!-- for (sample in grep('PBMC',names(pbmc.T),value=TRUE)) { -->
<!--   sobj <- pbmc.T[[sample]] -->
<!--   sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map) -->
<!--   sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))), -->
<!--                            levels=c('small','medium','large','hyperexpanded')) -->
<!--   tmp[[sample]] <- occupiedscRepertoire(sobj, -->
<!--                                         x.axis = "celltype",  -->
<!--                                         exportTable=TRUE) %>% -->
<!--     dplyr::mutate(sample=gsub('_PBMC','',sample)) -->
<!-- } -->

<!-- ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) +  -->
<!--   geom_col() +  -->
<!--   theme_classic() + -->
<!--   facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') + -->
<!--   scale_fill_manual(values=cloneType_colors) + -->
<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5), -->
<!--         strip.background=element_blank(), -->
<!--         strip.placement='outside', -->
<!--         legend.key.size=unit(3,'mm')) + -->
<!--   labs(x='',y='') -->
<!-- ``` -->

look at clonal diversity (Inv. Simpson) for different CD4 and CD8 subsets in PBMC separately, using the more fine-grained annotation

```{r Fig_S4C,fig.width=6,fig.height=2}
cols <- c('sample','orig.ident','predicted.celltype.l2')
all_meta <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) x@meta.data[,cols])) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list()
for (ct in grep('CD[48]|Treg',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  tcrcomb <- lapply(tcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
  dfs[[ct]] <- clonalDiversity(tcrcomb[lapply(tcrcomb,function(x) dim(x)[1]) > 0],
                               cloneCall='gene',group='sample',exportTable=TRUE, skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint =ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>% 
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
         dplyr::filter(metric=='Inv.Simpson') %>%
         dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
       aes(x=Timepoint,y=value)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(. ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(aes(group=Patient.ID),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format",paired=TRUE,size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_blank(),
        strip.background=element_blank(),
        strip.text=element_text(size=8),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip-t", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl
```

```{r Fig_S11F,fig.width=6,fig.height=3}
pl <- ggplot(df %>%
               dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>%
               gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
               dplyr::filter(metric=='Inv.Simpson') %>%
               dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
             aes(x=Responder,y=value)) +
  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=responder_colors,limits=force) +
  scale_color_manual(values=responder_colors,limits=force) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_blank(),
        strip.background=element_blank(),
        strip.text=element_text(size=8),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip-t", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl
```

check TCR diversity for CD4 / CD8 subsets that are

- ITGA4, ITGB7 or ITGB1 positive
- Integrin-b7 positive (after determining a threshold)
- GPR15,  CCR9 or ITGAE positive 
- ITGA4, ITGB7 or ITGB1 negative
- Integrin-b7 negative

```{r Fig_S4E_S4H_S5C_S5D,fig.width=2.5,fig.height=2.5, cache=FALSE}
library(Binarize)
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
}

cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','ITGB1','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

threshold <- binarize.kMeans(pbmc_md$`origadt_Integrin-b7`)@threshold
pbmc_md$Integrinb7 <- ifelse(pbmc_md$`origadt_Integrin-b7` > threshold,'pos','neg')

dfs <- list()
for (ct in c('Mmry CD4','Mmry CD8')) {
  bcs <- list(ITGA4=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4 > 0)],
              ITGB7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7 > 0)],
              Integrin_b7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='pos')],
              GPR15=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$GPR15 > 0)],
              CCR9=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$CCR9 > 0)],
              ITGAE=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGAE > 0)],
              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)],
              ITGB1=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1 > 0)],
              ITGA4neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4==0)],
              ITGB7neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7==0)],
              ITGB1neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1==0)],
              Integrin_b7_neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='neg')])
  for (nn in names(bcs)) {
    if (length(bcs[[nn]]) > 0) {
      tcrcomb <- lapply(tcr_combined,
                        function(x) x[(x$barcode %in% bcs[[nn]]) & !is.na(x$TCR1) & !is.na(x$TCR2),])
      dfs[[paste0(ct,'_',nn)]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
        dplyr::mutate(celltype=ct,
                      subset=nn)
    }
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

for (ss in c('GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7','ITGB1','ITGA4neg','ITGB7neg','ITGB1neg','Integrin_b7_neg')) {
  for (ct in c('Mmry CD4','Mmry CD8')) {
    pl <- ggplot(df %>% 
                   dplyr::filter(Diagnosis=='UC') %>%
                   gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,subset)) %>%
                   dplyr::filter(metric=='Inv.Simpson') %>%
                   dplyr::group_by(celltype,subset,Patient.ID) %>%
                   dplyr::mutate(ns=n()) %>%
                   dplyr::filter(ns==2,subset==ss,celltype==ct), 
                 aes(x=Timepoint,y=value)) +
      geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
      geom_line(aes(group=Patient.ID),color='gray',size=.25) +
      geom_point(size=1,color='gray30') +
      theme_classic() +
      coord_cartesian(clip='off') +
      scale_y_log10() +
      stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
      scale_fill_manual(values=timepoint_colors) +
      scale_color_manual(values=timepoint_colors) +
      labs(x='',y='Inv. Simpson',title=paste0(ct,'\n',ss)) +
      theme(legend.key.size=unit(4,'mm'),
            legend.box='horizontal',
            axis.text.x=element_text(angle=90,hjust=1,vjust=0),
            strip.background=element_blank(),
            strip.placement='outside')
    print(pl)
  }
}
```

```{r Fig_S4H,fig.width=2.5,fig.height=2}
ADT %>% 
  dplyr::group_by(orig.ident,Patient.ID, celltype,Timepoint) %>% 
  dplyr::summarise(pos=mean((expression > threshold))) %>%
  dplyr::filter(Timepoint %in% c("Vedo_0d","Vedo_6wk")) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2, celltype=='Mmry CD4+ T cells') %>%
  ggplot(aes(x=Timepoint,y=pos,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  #facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=90,hjust=1,vjust=.5))  +
  labs(x='',y='fraction Integrin-b7-positive cells',fill='',color='')
```

```{r Fig_S4G, fig.width=2.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype=='Mmry CD4+ T cells') %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  #facet_grid(.~celltype) +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='',title='Mmry CD4+ T cells')
```

```{r Fig_S5E, fig.width=2.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype=='Mmry CD8+ T cells') %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='',title='Mmry CD8+ T cells')
```

```{r Fig_S4D,fig.width=3,fig.height=2.5}
cols <- c('sample','orig.ident','predicted.celltype.l2')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=predicted.celltype.l2,
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list() 
#smps <- unique(gsub('_[CD4PMBC]*$','',names(tcr_combined)))
smps <- unique(gsub('_[CD4PMBC]*$','',grep('PBMC',names(tcr_combined),value=TRUE)))
for (ct in c('CD4 Naive','CD4 Proliferating','CD4 TCM','CD4 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct)]
  tcrcomb <- lapply(smps,
                    function(x) tcr_combined[[paste0(x,'_PBMC')]] %>%
                      dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
                      dplyr::mutate(sample=gsub('_PBMC$','',sample)))
  #                          tcr_combined[[paste0(x,'_CD4')]] %>%
  #                            dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
  #                            dplyr::mutate(sample=gsub('_CD4$tcr_','',sample)))),
  # lapply(setdiff(smps,smps_2),
  #        function(x) tcr_combined[[paste0(x,'_PBMC')]] %>%
  #          dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
  #          dplyr::mutate(sample=gsub('_PBMC$','',sample))))
  names(tcrcomb) <- smps #c(smps_2,setdiff(smps,smps_2))

  dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  #dplyr::mutate(sample=ifelse(grepl('_CD4$',sample),sample,paste0(sample,'_PBMC'))) %>%
  dplyr::mutate(sample=paste0(sample,'_PBMC')) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>%
               dplyr::filter(celltype %in% c('CD4 TCM','CD4 TEM'),
                             Timepoint!='ctrl')%>%
               dplyr::group_by(celltype,Diagnosis,Patient.ID) %>%
               dplyr::mutate(ns=n()) %>%
               dplyr::filter((ns > 1) | (Diagnosis=='control')),
             aes(x=Timepoint,y=`Inv.Simpson`)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(. ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  stat_compare_means(comparisons=list(c('ctrl','Vedo_0d')),label = "p.format",size=2,bracket.size=.25,tip.length=.01,step.increase=1) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip|panel|axis", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
```

```{r Fig_S4D_S5B,fig.width=4,fig.height=4}
dfs <- list()
for (ct in c('CD4 TCM','CD4 TEM',
             'CD8 TCM','CD8 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct) & (pbmc_md$Material=='PBMC')]
  if (length(bcs) > 0) {
    tcrcomb <- lapply(tcr_combined,
                      function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
    dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
      dplyr::mutate(celltype=gsub('CD[48] ','',ct),
                    Material=ifelse(grepl('CD4',ct),'CD4','CD8'))
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>%
               dplyr::group_by(celltype,Diagnosis,Material,Patient.ID) %>%
               dplyr::mutate(ns=n()) %>%
               dplyr::filter((ns==2) | (Diagnosis=='control')),
             aes(x=Timepoint,y=`Inv.Simpson`)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(width=.25,size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Material ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  stat_compare_means(comparisons=list(c('ctrl','Vedo_0d')),label = "p.format",size=2,bracket.size=.25,tip.length=.01,step.increase=1) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip|panel|axis", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)

#pl
```

```{r Fig_S5F,fig.width=4.5,fig.height=3.5,cache=FALSE}
pbmc.B <- lapply(names(bcr_combined), function(sample) {
  bcs <- bcr_combined[[sample]]$barcode
  cells <- paste0(sample,'_',sample,'_',Cells(pbmc[[sample]]))
  tmp <- RenameCells(pbmc[[sample]], new.names=cells)
  combineExpression(bcr_combined[[sample]][bcs %in% cells,],tmp, cloneCall="aa")
  })
names(pbmc.B) <- names(bcr_combined)

pbmc_B_meta <- do.call(rbind,
                       lapply(pbmc.B,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
  #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_B_meta %>% dplyr::mutate(BCR_detection=ifelse(is.na(CTaa),'no BCR','BCR')) %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=BCR_detection)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=c('BCR'='lightgreen','no BCR'='grey'),limits=force) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))

ggplot(pbmc_B_meta %>% dplyr::mutate(cloneType=factor(tolower(gsub('Single','Small',gsub(' .*','',cloneType))),
                                                      levels=c('small','medium','large','hyperexpanded'))) %>% dplyr::sample_frac(1L),
       aes(x=refUMAP_1,y=refUMAP_2,color=cloneType)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=cloneType_colors,na.value='gray80') +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

look at clonal diversity for B cells

```{r Fig_S5G,fig.width=4.5,fig.height=3}
dfs <- list()
for (ct in grep('B|Plasmablast',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  bcrcomb <- lapply(bcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !grepl('_NA|NA_',x$CTaa),])
  dfs[[ct]] <- clonalDiversity(bcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct,
                  Material='PBMC')
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

ggplot(df %>% 
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
         dplyr::filter(metric=='Inv.Simpson') %>%
         dplyr::group_by(Patient.ID,celltype) %>%
         dplyr::mutate(ns=n()) %>%
         dplyr::filter(ns==2), 
       aes(x=Timepoint,y=value)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(~celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')
```

```{r Fig_S10F,fig.width=4.5,fig.height=3}
ggplot(df %>%
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
         dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>%
         dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')),
       aes(x=Responder,y=value)) +
  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=responder_colors) +
  scale_color_manual(values=responder_colors) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')
```

<!-- ```{r clonotype_celltypes_BCR,fig.width=6.5,fig.height=10} -->
<!-- tmp <- list() -->
<!-- for (sample in grep('PBMC',names(pbmc.B),value=TRUE)) { -->
<!--   sobj <- pbmc.B[[sample]] -->
<!--   sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map) -->
<!--   sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))), -->
<!--                            levels=c('small','medium','large','hyperexpanded')) -->
<!--   tmp[[sample]] <- occupiedscRepertoire(sobj, -->
<!--                                         x.axis = "celltype", -->
<!--                                         exportTable=TRUE) %>% -->
<!--     dplyr::mutate(sample=gsub('_PBMC','',sample)) -->
<!-- } -->

<!-- ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) + -->
<!--   geom_col() + -->
<!--   theme_classic() + -->
<!--   facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') + -->
<!--   scale_fill_manual(values=cloneType_colors) + -->
<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5), -->
<!--         strip.background=element_blank(), -->
<!--         strip.placement='outside', -->
<!--         legend.key.size=unit(3,'mm')) + -->
<!--   labs(x='',y='') -->
<!-- ``` -->

```{r Fig_S5H, fig.width=4.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype %in% c('IgD- B cells','IgD+ B cells','Plasmablast')) %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  facet_grid(.~celltype) +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,hjust=.5,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='')
```

<!-- # Fig. 4 supplements -->

```{r get_GSE73661}
library(limma)
library(SummarizedExperiment)
olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11',
                 'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA',
                 'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF',
                 'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8',
                 'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1')

meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>%
  dplyr::mutate(sample=make.names(geo_accession)) %>%
  tibble::rownames_to_column("accession")  %>%
  dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')),
                treatment=factor(ifelse(diagnosis=='control','control',
                                        ifelse(grepl('plac',sample),'placebo',
                                               ifelse(grepl('IFX',sample),'IFX',
                                                      ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))),
                                 levels=c('control','placebo','IFX','vdz4w','vdz8w')),
                timepoint=factor(ifelse(diagnosis=='control','control',
                                        ifelse(grepl('W0',sample),'w0',
                                               ifelse(grepl('W4|W6',sample),'w6',
                                                      ifelse(grepl('W12',sample),'w12',
                                                             ifelse(grepl('W52',sample),'w52','other'))))),
                                 levels=c('control','other','w0','w6','w12','w52')),
                responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>%
  dplyr::filter(!grepl('earlier.stop',sample),
                !grepl('_47$',sample),
                !grepl('_73$',sample),
                (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>%
  droplevels()

row.names(meta) <- meta$sample
colnames(meta) <- gsub('.ch1','',colnames(meta))
data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t')
data <- data[,colnames(data) %in% row.names(meta)]

GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),])
```

show a heatmap of gene expression

```{r Fig_S17C,fig.width=10,fig.height=10}
olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12'))

ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')],
                        show_legend=TRUE,
                        col=list('timepoint'=c('control'='#000000',
                                               'w0'='#ff7700',
                                               'w6'='#882e72'),
                                 'responder'=c('control'='#000000',
                                               'R'='#db2F26',
                                               'NR'='#34a048'),
                                 'diagnosis'=c('UC'='#1965b0',
                                               'control'='#000000')))

mat <- t(scale(t(assay(GSE73661)[olink_genes_here,])))
Heatmap(mat,
        cluster_columns=TRUE,
        show_row_names=TRUE,
        show_column_names=FALSE,
        top_annotation=ha,
        column_names_gp=gpar(fontsize=6),
        row_names_gp=gpar(fontsize=6),
        name='z-score')
```

run a few contrasts:

- w6vsw0 @ UC
- UCvscontrol

```{r GSE73661_DE,fig.width=6,fig.height=5}
library(limma)
res <- list()

take <- colData(GSE73661)$diagnosis=='UC'
design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,])
fit <- eBayes(lmFit(assay(GSE73661)[,take], design))
contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design)
contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>%
  tibble::rownames_to_column('gene') %>%
  dplyr::filter(gene %in% olink_genes)  %>%
  dplyr::mutate(contrast='w6vsw0@UC')

design <- model.matrix(~0+diagnosis,data=colData(GSE73661))
fit <- eBayes(lmFit(assay(GSE73661), design))
contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design)
contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>%
  tibble::rownames_to_column('gene') %>%
  dplyr::filter(gene %in% olink_genes)  %>%
  dplyr::mutate(contrast='UCvscontrol')

GSE73661_DE <- do.call(rbind,res) %>%
    dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast)
```

show volcano plots (only for Olink genes!)

```{r Fig_S17A_B,fig.width=4.5,fig.height=4}
for (ct in c('w6vsw0@UC','UCvscontrol')) {
  res <- GSE73661_DE %>% dplyr::filter(contrast==ct)
  pl <-  res %>%
    ggplot(aes(x=logFC,y=-log10(P.Value))) +
    geom_point(aes(color=(P.Value < .05)),size=.5) +
    geom_text_repel(data=res %>% dplyr::filter(P.Value < .05),
                    aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') +
    theme_classic() +
    geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) +
    geom_vline(xintercept=0,linetype='dashed',size=.25) +
    scale_color_manual(values=c('black','red'),guide=FALSE) +
    labs(title=ct)
  print(pl)
}
```

```{r Fig_2A,fig.width=4,fig.height=3}
DimPlot(CD4.integrated,label=TRUE) +
  scale_color_brewer(palette='Set3') +
  theme_void() +
  theme(legend.key.size=unit(3,'mm'))
```

```{r Fig_2B,fig.width=4,fig.height=3,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'
FetchData(CD4.integrated, c('UMAP_1','UMAP_2','MKI67')) %>%
  dplyr::arrange(MKI67) %>%
  ggplot(aes(x=UMAP_1,y=UMAP_2,color=MKI67)) +
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       oob=scales::squish,
                       guide=guide_colorbar(title.position='right',
                                            title.theme=element_text(angle=90,hjust=.5,vjust=.5))) + 
  theme(legend.key.size=unit(3.5,'mm')) +
  labs(color='MKI67')
```

```{r CD4_umap_material,fig.width=4,fig.height=3}
DimPlot(CD4.integrated,group.by='Material') +
  scale_color_brewer(palette='Set1') +
  theme_void() +
  theme(legend.key.size=unit(3,'mm')) +
  labs(title='')
```

```{r CD4_umap_patient,fig.width=4,fig.height=3}
DimPlot(CD4.integrated,group.by='Patient.ID') +
  theme_void() +
  theme(legend.key.size=unit(3,'mm')) +
  labs(title='')
```

plot the cluster composition

```{r CD4_cluster_composition,fig.width=4,fig.height=3}
CD4.integrated[['cluster']] <- Idents(CD4.integrated)
for (col in c('Diagnosis','Patient.ID','Material')) {
  CD4.integrated$tmp <- CD4.integrated@meta.data[,col]
  pl <- rbind(CD4.integrated@meta.data %>% 
                dplyr::group_by(cluster,tmp) %>%
                dplyr::summarise(n=n()) %>%
                dplyr::mutate(frac=n/sum(n),
                              group='by cluster'),
              CD4.integrated@meta.data %>% 
                dplyr::group_by(tmp) %>%
                dplyr::summarise(n=n()) %>%
                dplyr::mutate(frac=n/sum(n),
                              group='all',
                              cluster='all')) %>%
    dplyr::mutate(cluster=factor(cluster,levels=c('all',seq(1,12)))) %>%
    ggplot(aes(x=cluster,y=frac,fill=tmp)) +
    geom_col(outlier.shape=NA) +
    theme_classic() +
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
          legend.key.size=unit(3,'mm'),
          strip.background=element_blank()) +
    facet_grid(.~group,scales='free_x',space='free_x') +
    labs(x='',y='fraction cells',fill='')
  print(pl)
}
```

```{r CD4_markers,fig.width=14,fig.height=4.5,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'
if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) {
  markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25,
                            thresh.use = 0.25,
                            print.bar=FALSE, verbose=FALSE)
  write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv'))
}

markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv'))

top10 <- markers %>%
  group_by(cluster) %>%
  top_n(10, avg_log2FC)

DotPlot(CD4.integrated, features = unique(top10$gene)) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.key.size=unit(3,'mm')) +
  labs(x='',y='cluster')
```

```{r Fig_2D, fig.width=2,fig.height=1,cache=FALSE}
library(gtools)
CD4.integrated[['cluster']] <- Idents(CD4.integrated)

cols <- c('sample','orig.ident','cluster','Patient.ID','Responder','Diagnosis','Material','Timepoint','Batch')
CD4_frac <- CD4.integrated@meta.data[,cols] %>%
  dplyr::mutate(celltype=cluster) %>%
  dplyr::group_by(sample,Patient.ID,Responder,Diagnosis,Material,Timepoint,Batch,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n))

CD4_frac$Timepoint <- factor(CD4_frac$Timepoint, levels=c('ctrl','Vedo_0d','Vedo_6wk'))
CD4_frac$Responder <- revalue(CD4_frac$Responder, c('yes'='R','no'='NR'))

CD4_frac %>%
  dplyr::filter(Timepoint!='ctrl',celltype=='1',Timepoint=='Vedo_0d') %>%
  ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
  geom_boxplot(outlier.shape=NA) +
  #geom_point(size=1,aes(shape=Material),color='gray30') +
  geom_point(size=1,color='gray30') +
  stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  coord_flip() +
  scale_x_discrete(limits=rev) +
  scale_fill_manual(values=responder_colors,limits=force) +
  scale_color_manual(values=responder_colors,limits=force) +
  scale_shape_manual(values=c(1,4)) +
  theme(legend.position='none',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside')  +
  labs(x='',y='cell fraction',fill='',color='',shape='')
```

<!-- ```{r Fig_S19E, fig.width=5,fig.height=2.5} -->
<!-- CD4_frac %>% -->
<!--     dplyr::filter(Timepoint!='ctrl', celltype %in% seq(1,6)) %>% -->
<!--     ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) +  -->
<!--     geom_boxplot(outlier.shape=NA) + -->
<!--     geom_point(size=1,aes(shape=Material),color='gray30') + -->
<!--     stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format",  -->
<!--                        method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) + -->
<!--     theme_classic() +  -->
<!--     coord_cartesian(clip='off') + -->
<!--     facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
<!--     scale_fill_manual(values=responder_colors,limits=force) + -->
<!--     scale_color_manual(values=responder_colors,limits=force) + -->
<!--     scale_shape_manual(values=c(1,4)) + -->
<!--     theme(axis.text.x=element_blank(), -->
<!--           legend.position='right', -->
<!--           legend.key.size=unit(3.5,'mm'), -->
<!--           strip.background=element_blank(), -->
<!--           strip.placement='outside')  + -->
<!--     labs(x='',y='cell fraction',fill='',color='',shape='') -->

<!-- CD4_frac %>% -->
<!--     dplyr::filter(Timepoint!='ctrl', celltype %in% seq(7,12)) %>% -->
<!--     ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) +  -->
<!--     geom_boxplot(outlier.shape=NA) + -->
<!--     geom_point(size=1,aes(shape=Material),color='gray30') + -->
<!--     stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format",  -->
<!--                        method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) + -->
<!--     theme_classic() +  -->
<!--     coord_cartesian(clip='off') + -->
<!--     facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
<!--     scale_fill_manual(values=responder_colors,limits=force) + -->
<!--     scale_color_manual(values=responder_colors,limits=force) + -->
<!--     scale_shape_manual(values=c(1,4)) + -->
<!--     theme(axis.text.x=element_blank(), -->
<!--           legend.position='right', -->
<!--           legend.key.size=unit(3.5,'mm'), -->
<!--           strip.background=element_blank(), -->
<!--           strip.placement='outside')  + -->
<!--     labs(x='',y='cell fraction',fill='',color='',shape='') -->
<!-- ``` -->

```{r Fig_2C, fig.width=3,fig.height=2,cache=FALSE}
library(gtools)
CD4.integrated[['frac_MKI67']] <- ifelse(CD4.integrated@assays$RNA@counts['MKI67',] > 0,'MKI67+','MKI67-')

cols <- c('sample','orig.ident','cluster','Patient.ID','Responder','Diagnosis','Material','Timepoint','Batch','frac_MKI67')
CD4_frac <- CD4.integrated@meta.data[,cols] %>%
  dplyr::mutate(celltype=cluster) %>%
  dplyr::group_by(sample,Patient.ID,Responder,Diagnosis,Material,Timepoint,Batch,celltype) %>%
  dplyr::summarise(frac=mean(frac_MKI67=='MKI67+'))

CD4_frac %>%
  dplyr::filter(Timepoint!='ctrl') %>%
  ggplot(aes(x=celltype,y=frac,fill=celltype,color=celltype)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  theme_classic() + 
  coord_cartesian(clip='off') +
  scale_fill_brewer(palette='Set3') +
  scale_color_brewer(palette='Set3') +
  scale_shape_manual(values=c(1,4)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside')  +
  labs(x='',y='fraction MKI67+',fill='',color='')
```

```{r CD4_regression_DE,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'
CD4.integrated$cluster1 <- factor(ifelse(Idents(CD4.integrated)=='1','1','other'),levels=c('1','other'))

genes <- list('activation'=c('HLA-DRA','CD40LG','CD38','CD69','CTLA4','AREG','ICOS','LAG3','MKI67','PDCD1','SELL','SLAMF6','HAVCR2','TIGIT'),
              'homing'=c('CXCR3','CXCR5','CCR2','CCR4','CCR5','CCR6','CCR7','CCR9','GPR15','ITGA4','ITGB1','ITGB7'),
              'cytokines'=c('CSF2','IFNG','IL2','IL4','IL5','IL9','IL10','IL17A','IL22','TNF','GZMA','GZMB'),
              'receptors'=c('IL1R1','IL1RAP','IL2RA','IL4R','IL7R','IL12RB2','IL18RAP','IL18R1','IL23R','IFNGR1','IFNGR2','TNFRSF1A','TNFRSF1B','TNFRSF14'),
              'TFs'=c('AHR','BATF','BCL6','CREM','EOMES','FOSB','GATA3','IKZF2','JUN','JUNB','MAF','NFATC1','PRDM1','RORA','RORC','TCF7','TOX','ZBTB7B','TBX21'))

if (!file.exists(file.path(path,'seurat','CD4_cluster1_markers.csv'))) { 
  CD4_cluster1_markers <- FindMarkers(CD4.integrated,test.use='LR',logfc.threshold=0,
                                      latent.vars=c('Material'),min.pct=0.0,
                                      group.by='cluster1',ident.1='1',ident.2='other')
#                                      features=intersect(rownames(CD4.integrated),unname(unlist(genes))))
  write.csv(CD4_cluster1_markers,file.path(path,'seurat','CD4_cluster1_markers.csv')) 
} else {
  CD4_cluster1_markers <- read.csv(file.path(path,'seurat','CD4_cluster1_markers.csv'),row.names=1)
}
```

as scatter plot

```{r Fig_2E,fig.width=4,fig.height=3.5}
top_genes <- CD4_cluster1_markers %>% 
  tibble::rownames_to_column('gene') %>%
  dplyr::slice_max(abs(avg_log2FC),n=30) %>%
  dplyr::pull(gene)
top_genes <- setdiff(c(top_genes,c('ITGA4','ITGB7','CD38','MKI67','CXCR3','CCR7')),c('C12orf75','FLNA','HLA−DRB5','CD99','YWHAQ','HLA−DPB1'))
ggplot(CD4_cluster1_markers,
  aes(x=pct.2,y=pct.1)) +
  geom_abline(slope=1,intercept=0,size=.5,color='gray') +
  #geom_point(size=1,pch=20,alpha=.5,color='gray') +
  geom_point(alpha=.5,size=1,color='gray30') +
  geom_label_repel(data=CD4_cluster1_markers[top_genes,] %>%
                     tibble::rownames_to_column('gene'),
                  aes(label=gene), force=5,max.overlaps=25,
                  label.padding=unit(.5,'mm'),
                  label.size=0,
                  color='black',
                  size=2.5,segment.alpha=.5,segment.size=.5) +
  scale_color_gradient2(low='blue3',mid='gray',high='red3') +
  labs(x='% positive cells other clusters',
       y='% positive cells cluster 1',
       fill='log2 fold change',
       size='p-value') +
  theme_classic() +
  #coord_cartesian(xlim=c(-.1,1),ylim=c(-.1,1)) +
  #guides(color=guide_legend(override.aes=list(shape=15,size=3,label="o"))) +
  scale_shape_manual(values=seq(21,26)) +
  theme(legend.key.size=unit(3,'mm'))
```

as lollipop chart

```{r Fig_2F,fig.width=10,fig.height=3}
CD4_cluster1_markers %>% 
  tibble::rownames_to_column('gene') %>%
  dplyr::inner_join(data.frame(gene=unlist(genes)) %>% tibble::rownames_to_column('group') %>% dplyr::mutate(group=gsub('[0-9]*$','',group)),
                    by='gene') %>%
  dplyr::mutate(group=factor(group),
                gene=factor(gene,levels=unname(unlist(genes)))) %>%
  dplyr::arrange(group,avg_log2FC) %>%
  ggplot(aes(x=gene,y=avg_log2FC,color=avg_log2FC)) +
  geom_segment(aes(xend=gene,yend=0)) +
  geom_point(aes(size=-log10(p_val+1.e-30)),pch=19) +
  scale_color_gradient2(low='blue3',high='red3') +
  labs(y='avg. log2 fold change',
       x='',
       fill='log2 fold change',
       size='p-value') +
  facet_grid(.~group,scales='free_x',space='free_x') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_size_continuous(limits=c(0,50),breaks=c(10,20,30),labels=c('1.e-10','1.e-20','<1.e-30')) +
  theme(legend.key.size=unit(3,'mm'),
        axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        strip.background=element_blank(),
        strip.clip='off')
```

```{r Fig_S8D,fig.width=4,fig.height=3.5,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'orig_ADT'
DotPlot(CD4.integrated, features = c('PD-1','CD25','CD45RA','CD127','CD38','CD103','CD4','HLA-DR'),
        group.by='cluster1', scale=TRUE) + 
  scale_color_viridis_c() +
  scale_x_discrete(limits=rev) +
  coord_flip() +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.key.size=unit(3,'mm'),
        legend.box='vertical') +
  labs(x='',y='')
```

```{r CD4_clonotypes,fig.width=3,fig.height=2.5,cache=FALSE}
CD4.integrated <- RenameCells(CD4.integrated, add.cell.id=CD4.integrated$orig.ident)

tmp <- lapply(tcr_combined, function(x) x[x$barcode %in% Cells(CD4.integrated),])
CD4.integrated <- combineExpression(tmp, CD4.integrated,  cloneCall="aa")
CD4.integrated$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',CD4.integrated$cloneType))),
                                   levels=c('small','medium','large','hyperexpanded'))

CD4_md <- CD4.integrated@meta.data %>%
  tibble::rownames_to_column('cell')

dfs <- list()
for (cl in unique(CD4.integrated$cluster)) {
  bcs <- CD4_md$cell[CD4_md$cluster==cl]
  tcrcomb <- lapply(tcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
  dfs[[cl]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(cluster=cl)
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

ggplot(df %>% 
         #dplyr::filter(Diagnosis=='UC') %>%
         dplyr::mutate(cluster=factor(cluster,levels=as.character(seq(0,12)))), 
       aes(x=cluster,y=`Inv.Simpson`)) +
  geom_boxplot(aes(fill=cluster),outlier.shape=NA) +
  geom_point(width=.25,size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  labs(x='cluster',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        legend.position='none',
        axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        strip.background=element_blank(),
        strip.placement='outside')
```

<!-- ```{r tmp} -->
<!-- tmp <- CD4_md %>% dplyr::filter(!is.na(CTaa)) %>% dplyr::select(cluster,sample) %>% table() %>% as.data.frame()  -->
<!-- dplyr::inner_join(df,tmp,by=c('cluster','sample')) %>%  -->
<!--   ggplot(aes(x=Freq,y=`Inv.Pielou`,label=cluster,color=cluster=='1')) +  -->
<!--   geom_text() -->
<!-- ``` -->
```{r sessionInfo}
sessionInfo()
```