### 2024-09-25 ML_Response_prediction_visualization.rmd
20240925 Updated scripts for publication

---
title: "Visualization ML prediction"
output: html_notebook
---

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpattern)
library(ggpubr)

theme_figures <- function() {
  theme_minimal() %+replace%    #replace elements we want to change
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 8),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8),
          strip.text.x = element_text(size = 8)) }
```
# Model performances

## AUC values

```{r auc_read_in}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_response_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*pred_test.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]), row.names = 1))
  colnames(data) <- c('TrainingMeanAUC', 'TrainingStdAUC', 'TestingMeanAUC', 'TestingStdAUC')
  data <- t(round(data, 2))
  mat <- matrix(data, nrow = 2, ncol = 2)
  rownames(mat) <- c("MeanAUC", "SEMAUC")
  colnames(mat) <- c("Training", "Test")

  combined_data[[i]] <- mat
}
# Tidy dataframe
names(combined_data) <- gsub(pattern = "R_response_|_pred_test.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_accuracy <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "Set", "MeanAUC", 'SEMAUC')) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                       ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                              ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                     ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest")) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
  mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))
```

```{r auc_vis_single_modalities}
# Code Supplement 18 A
pdf("../PredictiveModels/results/S_18A.pdf", height = 2, width = 4)
p <- df_accuracy %>%
  filter(DataSet %in% c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS")) %>%
  filter(FeatureEngineering == "No") %>%
  mutate(DataSet = factor(DataSet, levels = c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS"))) %>%
  mutate(DataSet = gsub("only", replacement = "", DataSet)) %>%
  mutate(DataSet = gsub("CyTOFfreq", replacement = "CyTOF", DataSet)) %>%
  mutate(DataSet = gsub("CyTOFA4B7", replacement = "CyTOFab", DataSet)) %>%
  mutate(DataSet = factor(DataSet, levels = c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS"))) %>%
  mutate(Classifier = factor(Classifier, levels = c("Treatment", "Response"))) %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  # Data visualization -------------------------------------------
  ggplot(aes(x = DataSet, y = MeanAUC, fill = Method)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.6) +
  geom_errorbar(aes(ymin = MeanAUC - SEMAUC, ymax = MeanAUC + SEMAUC), width = .2,
                position = position_dodge(.6)) +
  geom_hline(yintercept = 0.8, linetype = "dashed",
             color = "black") +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "red") +
  scale_fill_manual(values = c("#DDDDDD", "#AEADAD", "#7A7A7A", "#505050", "#1B1B1B")) +
  theme_figures() +
  ylab("AUC") +
  coord_cartesian(ylim = c(0, 1))
print(p)
dev.off()
# #-------------------------------------------------------------------------------------
print(p)
```

```{r}
good_models <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(SelectionCriteria >= 0.5) %>%
  filter(FeatureEngineering == "No") %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  pull(DataSet)

# Code Supplement 18 B
pdf("../PredictiveModels/results/S_18B.pdf", height = 2.15, width = 8.1)
p <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(DataSet %in% good_models) %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>%
  mutate(FeatureCutoff = factor(FeatureCutoff, levels = c("none", "95%", "90%", "80%", "75%", "50%"))) %>%
  mutate(DataSet = factor(DataSet, levels = c("OLINK", "FACS", "OLINK\nClinical", "OLINK\nFACS", "FACS\nCyTOF", "FACS\nCyTOFab", "FACS\nCyTOF\nClinical", "FACS\nCyTOFab\nClinical", "FACS\nCyTOF\nOLINK", "FACS\nClinical\nOLINK", "all\nCyTOF"))) %>%
  ggplot(aes(x = DataSet, y = MeanAUC, fill = FeatureCutoff)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.6) +
  geom_errorbar(aes(ymin = MeanAUC - SEMAUC, ymax = MeanAUC + SEMAUC), width = .2,
                position = position_dodge(.6)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "red") +
  geom_hline(yintercept = 0.8, linetype = "dashed",
             color = "black") +
  scale_fill_manual(values = c("#000000", "#3B3B3B", "#636363", "#929292", "#C3C3C3", "#DFDFDF")) +
  theme_figures() +
  ylab("AUC") +
  coord_cartesian(ylim = c(0, 1)) +
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(legend.key.size = unit(0.5, 'lines'))
print(p)
dev.off()

print(p)
```

```{r}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_response_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*_coefficients.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                       ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                              ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                     ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest")) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
  mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))
```

```{r}
# Code Supplement 18D
pdf("../PredictiveModels/results/S_18C.pdf", height = 3.7, width = 4.8)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% good_models) %>%
  filter(RedPatientSet == 'No') %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>%
  mutate(DataSet = factor(DataSet, levels = c("FACS", "OLINK\nFACS", "FACS\nCyTOF", "FACS\nCyTOFab", "OLINK\nClinical", "FACS\nCyTOF\nClinical", "FACS\nCyTOFab\nClinical", "FACS\nCyTOF\nOLINK", "FACS\nClinical\nOLINK", "all\nCyTOF"))) %>%
  mutate(Feature = factor(Feature, levels = c("Ki67_FACS", "GPR15_FACS", "CXCR3_FACS", "IL4_FACS", "CCR7_FACS", "IFNg_FACS", "Treg_FACS", "B7_FACS", "IL17A_FACS", "HLADR_FACS", 'Tbet_FACS', "CD103_FACS", "IL.10RA", "MCP.3", "TNF", "NT.3", "CCL25", "TNFB", "CASP.8", "AXIN1", "OSM", "CRP", "IL8", "NK.cells"))) %>%
  group_by(Feature, DataSet) %>%
  dplyr::count(Feature, DataSet) %>%
  mutate(n_norm = n / 10) %>%
  ggplot(aes(x = Feature, y = DataSet)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8)) +
  coord_flip()
dev.off()
```