### 2024-10-14 machine learning/ML_Iterative_model_validation_StabilitySelection.rmd
Add files via upload
Added ML code into desired folder.

---
title: "Visualizations of Classification pipeline"
author: "Lisa M. Steinheuer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  This script visualizes the output of the python pipeline to predict Vedolizumab treatment response in IBD patients. After the revision we decided to include LASSO and ELASTICNET to validate the feature selection process.
output:
  html_document:
    toc: true # table of content true
    toc_depth: 5  # upto five depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 4)
```

# Preface

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpubr)

project_name <- "Vedo_Stab"

theme_lisa <- function() {

  theme_minimal() %+replace%    #replace elements we want to change

    theme(axis.title.x = element_blank(),
          #axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 6),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 6),
          strip.text.x = element_text(size = 6)) }

color_vector <- c("Logistic Regression" = "#762a83",
                  "LASSO" = "#9970ab",
                  "ELASTICNET" = "#e7d4e8",
                  "Linear SVM" = "#d9f0d3",
                  "Polynomial SVM" = "#a6dba0",
                  "Radial SVM" = "#5aae61",
                  "Random Forest" = "#1b7837")

color_vector_paper <- c("Logistic Regression" = "#C20000",
                        #"LASSO" = "#9970ab",
                        #"ELASTICNET" = "#e7d4e8",
                        "Linear SVM" = "#cccccc",
                        "Polynomial SVM" = "#969696",
                        "Radial SVM" = "#636363",
                        "Random Forest" = "#252525")

cols_modality <- c("FACS" = "#c01c2c",
                   "OLINK" = "#2e3e9b",
                   "CyTOF" = "#5e2590",
                   "CyTOF a4b7" = "#ab43a8",
                   "Clinical" = "#656565")

cols_feature_cutoff <- c("none" = "#252525", "50%" = "#636363", "75%" = "#969696", "80%" = "#bdbdbd", "90%" = "#C20000", "95%" = "#d9d9d9")

check_robustness <- function(n_features, threshold, reference_bootstrap, which_modality) {
  robust_features <- df_features %>%
    filter(DataSet == that_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == "Logistic Regression") %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
    mutate(absCoef = abs(Coefficient)) %>%
    group_by(DataSet, Feature, Run) %>%
    summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
    mutate(adjusted_mean = mean_features - sem_features) %>%
    arrange(DataSet, desc(adjusted_mean)) %>%
    group_by(DataSet, Run) %>%
    top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
    mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
    mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
    group_by(Feature) %>%
    #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
    filter(Bootstrap == reference_bootstrap) %>%
    filter(rank_per_run <= n_features) %>%
    pull(Feature)

  in_how_many_runs <- df_features %>%
    filter(DataSet == that_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == "Logistic Regression") %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
    mutate(absCoef = abs(Coefficient)) %>%
    group_by(DataSet, Feature, Run) %>%
    summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
    mutate(adjusted_mean = mean_features - sem_features) %>%
    arrange(DataSet, desc(adjusted_mean)) %>%
    group_by(DataSet, Run) %>%
    top_n(n = round(n() * threshold), wt = adjusted_mean) %>%
    mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
    mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
    filter(Bootstrap != reference_bootstrap) %>%
    group_by(Bootstrap) %>%
    summarise(contains_features = all(robust_features %in% Feature)) %>%
    summarise(count = sum(contains_features))
  output <- round(in_how_many_runs$count * 100 / 49, 2)
  return(output)

}

```

# Evaluation

## Model performances

```{r auc_read_in}

# Data read in
files_accurcy <- intersect(list.files(path = "/Iterative_model_validation/", pattern = "R_*"),
                           list.files(path = "/Iterative_model_validation/", pattern = "*_testing.csv"))

only_individual <- files_accurcy[grepl(pattern = "only", files_accurcy)]

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("/Iterative_model_validation/", files_accurcy[i]), row.names = 1))

  combined_data[[i]] <- data
}
# Data tweaking
names(combined_data) <- gsub(pattern = "R_|_testing.csv", replacement = "", files_accurcy)

df_testing_performance <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "remove_col", "Run", "Accuracy", 'F1')) %>%
  select(-remove_col) %>%
  mutate(Variable = gsub(Condition, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogisticRegression_", Condition), "Logistic Regression",
                                ifelse(grepl(pattern = "LASSO_", Condition), "LASSO",
                                       ifelse(grepl(pattern = "ELASTICNET_", Condition), "ELASTICNET",
                                              ifelse(grepl(pattern = "LinearSVM_", Condition), "Linear SVM",
                                                     ifelse(grepl(pattern = "PolynomialSVM_", Condition), "Polynomial SVM",
                                                            ifelse(grepl(pattern = "RadialSVM_", Condition), "Radial SVM",
                                                                   ifelse(grepl(pattern = "RandomForest_", Condition), "Random Forest", NA))))))), levels = c("Logistic Regression", "LASSO", "ELASTICNET", "Linear SVM", "Polynomial SVM", "Radial SVM", "Random Forest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(DataSet = gsub("^_", replacement = "", DataSet)) %>%
  mutate(Run = gsub(pattern = ".*_Run", replacement = "", x = DataSet)) %>%
  mutate(DataSet = gsub(pattern = "_Run[0-9]|_Run[0-9][0-9]", replacement = "", DataSet)) %>%
  mutate(Combination = ifelse(DataSet %in% c("Clinical", "CyTOF", "CyTOFa4b7", "OLINK", "FACS"), yes = "Single", no = "Combination")) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50)))

```

## The model coefficients

```{r prepare_coef}
files_accurcy <- intersect(list.files(path = "/Iterative_model_validation/", pattern = "R_*"),
                           list.files(path = "/Iterative_model_validation/", pattern = "*_coefficients.csv"))

only_individual <- files_accurcy[grepl(pattern = "only", files_accurcy)]

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("/Iterative_model_validation/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  #dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogisticRegression_", Condition), "Logistic Regression",
                                ifelse(grepl(pattern = "LASSO_", Condition), "LASSO",
                                       ifelse(grepl(pattern = "ELASTICNET_", Condition), "ELASTICNET",
                                              ifelse(grepl(pattern = "LinearSVM_", Condition), "Linear SVM",
                                                     ifelse(grepl(pattern = "PolynomialSVM_", Condition), "Polynomial SVM",
                                                            ifelse(grepl(pattern = "RadialSVM_", Condition), "Radial SVM",
                                                                   ifelse(grepl(pattern = "RandomForest_", Condition), "Random Forest", NA))))))), levels = c("Logistic Regression", "LASSO", "ELASTICNET", "Linear SVM", "Polynomial SVM", "Radial SVM", "Random Forest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(DataSet = gsub(DataSet, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(DataSet = gsub("^_", replacement = "", DataSet)) %>%
  mutate(Run = gsub(pattern = ".*_Run", replacement = "", x = DataSet)) %>%
  mutate(DataSet = gsub(pattern = "_Run[0-9]|_Run[0-9][0-9]", replacement = "", DataSet)) %>%
  mutate(Combination = ifelse(DataSet %in% c("Clinical", "CyTOF", "CyTOFa4b7", "OLINK", "FACS"), yes = "Single", no = "Combination"))
```

## Adding Origin

```{r fig.height = 3}
df_features <- df_features %>%
  mutate(Origin = factor(ifelse(grepl(patter = "_FACS", x = Feature), yes = "FACS", no =
    ifelse(grepl(patter = "_A4B7", x = Feature), yes = "CyTOF a4b7", no =
      ifelse(Feature %in% c("Basophils", "cDCs", "cMonocytes", "Eosinophils", "gd.T.cells", "Memory.B.cells", "Naive.B.cells", "MAIT.cells", "Memory.CD4", "Memory.CD8", "Naive.CD4", "Naive.CD8", "ncMonocytes", "NK.cells", "pDCs", "Plasmablasts"), yes = "CyTOF", no =
        ifelse(Feature %in% c('Leukocytes', 'CRP', 'Thrombocytes', 'Hemoglobin', 'Neutrophils', 'TNF_pretreated'), yes = "Clinical", no = "OLINK")))), levels = c("FACS", "OLINK", "CyTOF", "CyTOF a4b7", "Clinical")))

```

This code chunk checks how often a feature was included in the top 10 features from the non-feature selected data. I used the non-feature engineered data because otherwise I cannot extract the top 10 features per run. The smallest feature space was 22.
All in all, we have 18 model combinations.
I implemented it in a way, that we have for each non-responder (4) 4 different responder pairs

## Functions

```{r fig.width = 15, fig.height = 10}

Analyse_Modality_overview <- function(which_modality, which_classifier) {

  if (which_classifier %in% c("LASSO", "ELASTICNET")) {
    order_features <- df_features %>%
      filter(DataSet == that_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      filter(Coefficient != 0) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(Feature, Run) %>%
      dplyr::count(Feature) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      add_count(name = "Frequency") %>%
      arrange(-Frequency) %>%
      pull(Feature) %>%
      unique()

    p1 <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      filter(Coefficient != 0) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      mutate(Feature2 = factor(Feature, levels = order_features)) %>%
      ggplot(aes(x = Feature2, y = Bootstrap, fill = rank_per_run)) +
      geom_tile() +
      theme_minimal() +
      geom_text(aes(label = rank_per_run), size = 2.3) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      scale_fill_gradient("Rank", low = "#737373", high = "#f0f0f0", limits = c(0, 10)) +
      xlab("Features") +
      ylab("Within \n each iteration") +
      theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank()) +
      theme(legend.position = "none")

  } else if (which_classifier == "Logistic Regression") {
    order_features <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(DataSet, Run) %>%
      top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
      group_by(Feature, Run) %>%
      dplyr::count(Feature) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      add_count(name = "Frequency") %>%
      arrange(-Frequency) %>%
      pull(Feature) %>%
      unique()

    p1 <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(DataSet, Run) %>%
      top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
      mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      mutate(Feature2 = factor(Feature, levels = order_features)) %>%
      ggplot(aes(x = Feature2, y = Bootstrap, fill = rank_per_run)) +
      geom_tile() +
      theme_minimal() +
      geom_text(aes(label = rank_per_run), size = 2.3) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      scale_fill_gradient("Rank", low = "#737373", high = "#f0f0f0", limits = c(0, 10)) +
      xlab("Features") +
      ylab("Within \n each iteration") +
      theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank()) +
      theme(legend.position = "none") }

  p2 <- df_features %>%
    filter(DataSet == which_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == which_classifier) %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>%
    filter(Feature %in% order_features) %>%
    mutate(Feature2 = factor(Feature, levels = order_features)) %>%
    ggplot(aes(x = Feature2, y = 1, fill = Origin)) +
    geom_tile() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = cols_modality) +
    ylab("") +
    xlab("Feature") +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank()) +
    theme(legend.position = "bottom")

  p3 <- df_testing_performance %>%
    filter(DataSet == which_modality) %>%
    filter(Method == which_classifier) %>%
    filter(Combination == "Combination") %>%
    ggplot(aes(x = Method, y = Bootstrap, fill = Accuracy)) +
    geom_tile() +
    geom_text(aes(label = round(Accuracy, 1))) +
    scale_fill_gradient(low = "white", high = "#fc9272", limits = c(0, 1)) +
    theme_minimal() +
    theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank(),
          # axis.text.y = element_blank(),
          axis.ticks.y = element_blank())

  p4 <- cowplot::plot_grid(p1, p3, p2, labels = c('', '', ''), ncol = 2, rel_heights = c(0.8, 0.3), rel_widths = c(1, 0.15), align = 'v')

  ggsave(p4, file = paste0("../", project_name, "_Stability_Top10_", which_modality, "_", which_classifier, ".pdf"), width = 14, height = 9)

  print(p4)
  return }
```

```{r}
run_stability_analysis <- function(which_modality, rerun) {

  rerun_script <- rerun

  if (rerun_script == 1) {
    stability_landscape <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Stability" = integer())
    for (run in 1:50) {
      stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Stability" = integer())
      for (n_feat in 1:10) {
        for (cutoff in seq(0.1, 1.0, by = 0.1)) {
          how_many_features <- n_feat
          define_threshold <- cutoff
          define_reference <- run
          stability_value <- check_robustness(n_features = how_many_features, threshold = define_threshold, reference_bootstrap = define_reference, which_modality = that_modality)
          stability_landscape_int <- data.frame("Number of features" = how_many_features, "Threshold (top X%)" = define_threshold * 100, "Reference Run" = define_reference, "Stability" = stability_value)
          stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) } }

      print(mean(stability_landscape$Stability))

      stability_landscape <- rbind(stability_landscape, stability_landscape_iteration)
    }
    save(stability_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

  p1 <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    group_by(`Reference Run`) %>%
    summarise(`Mean Stability` = mean(Stability)) %>%
    ungroup() %>%
    ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean Stability`)) +
    geom_tile() +
    theme_minimal() +
    geom_text(aes(label = round(`Mean Stability`, 2)), size = 3.5) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    scale_fill_gradient("Stability", low = "#fee5d9", high = "#a50f15", limits = c(0, 100)) +
    scale_x_continuous(breaks = seq(1, 1, by = 1)) +
    scale_y_continuous(breaks = seq(0, 100, by = 10)) +
    ylab(paste("Iterations"))

  print(p1)
  ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

  refine_reference_run <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    group_by(`Reference Run`) %>%
    summarise(`Mean Stability` = mean(Stability)) %>%
    filter(`Mean Stability` == max(`Mean Stability`)) %>%
    slice(1)

  print(refine_reference_run)

  p2 <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
    ggplot(aes(x = `Number of features`, y = `Threshold (top X%)`, fill = Stability)) +
    geom_tile() +
    theme_minimal() +
    geom_text(aes(label = Stability), size = 2.5) +
    scale_fill_gradient("Stability", low = "#fee5d9", high = "#a50f15", limits = c(0, 100)) +
    scale_x_continuous(breaks = seq(1, 10, by = 1)) +
    scale_y_continuous(breaks = seq(10, 100, by = 10)) +
    xlab(paste("Number of features")) +
    labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
  ggsave(p2, file = paste0("../StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 4)
  data.frame(stability_landscape %>%
               set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
               filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
               arrange(desc(Stability), `Threshold (top X%)`) %>%
               filter(`Number of features` >= 5) %>%
               filter(`Threshold (top X%)` < 40))

  return(stability_landscape) }

```

## 4 best modalities

### OLINK_FACS

```{r fig.width = 15, fig.height = 10}
that_modality <- "OLINK_FACS"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)

number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("~/Desktop/", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_olink_facs)

feature_info_OLINK_FACS <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_olink_facs)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_olink_facs) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_olink_facs)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_olink_facs) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_olink_facs)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_olink_facs %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_olink_facs$Feature), feature_panel_olink_facs)

```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_olink_facs %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_olink_facs$Feature), feature_panel_olink_facs)
```

```{r}
summary_reg_olink_facs <- rbind(el_run_olink_facs, lasso_run_olink_facs) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### OLINK_FACS_CyTOF

```{r fig.width = 15, fig.height = 10}
that_modality <- "OLINK_FACS_CyTOF"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)

number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))
#

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  #geom_text(aes(label = AUC), size = 2.5) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  #scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("~/Desktop/", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

feature_info_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF)

print(feature_panel_OLINK_FACS_CyTOF)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_OLINK_FACS_CyTOF)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%

  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_OLINK_FACS_CyTOF)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_OLINK_FACS_CyTOF %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_OLINK_FACS_CyTOF$Feature), feature_panel_OLINK_FACS_CyTOF)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_OLINK_FACS_CyTOF %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_OLINK_FACS_CyTOF$Feature), feature_panel_OLINK_FACS_CyTOF)
```
```{r}
summary_reg_OLINK_FACS_CyTOF <- rbind(el_run_OLINK_FACS_CyTOF, lasso_run_OLINK_FACS_CyTOF) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### FACS_CyTOF_Clinical

```{r fig.width = 15, fig.height = 10}
that_modality <- "FACS_CyTOF_Clinical"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)
number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_FACS_CyTOF_Clinical)

feature_info_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_FACS_CyTOF_Clinical)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_info_FACS_CyTOF_Clinical) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOF_Clinical)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  #mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_FACS_CyTOF_Clinical) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOF_Clinical)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOF_Clinical %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_FACS_CyTOF_Clinical$Feature), feature_panel_FACS_CyTOF_Clinical)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOF_Clinical %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_FACS_CyTOF_Clinical$Feature), feature_panel_FACS_CyTOF_Clinical)
```

```{r}
summary_reg_FACS_CyTOF_Clinical <- rbind(el_run_FACS_CyTOF_Clinical, lasso_run_FACS_CyTOF_Clinical) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### FACS_CyTOFab

```{r fig.width = 15, fig.height = 10}
that_modality <- "FACS_CyTOFa4b7"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)
number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_FACS_CyTOFab)

feature_info_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_FACS_CyTOFab)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_info_FACS_CyTOFab) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOFab)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("~/Desktop/", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  #mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_FACS_CyTOFab) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOFab)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOFab %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_FACS_CyTOFab$Feature), feature_panel_FACS_CyTOFab)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOFab %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_FACS_CyTOFab$Feature), feature_panel_FACS_CyTOFab)

summary_reg_FACS_CyTOFab <- rbind(el_run_FACS_CyTOFab, lasso_run_FACS_CyTOFab) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))

```

```{r}
order_of_features <- c("Ki67", "CXCR3", "IL4", "GPR15", "CD103", "IFNg","B7", "CXCL6", "TNF", "NK cells")

combined_regularizers <- rbind(summary_reg_olink_facs, summary_reg_FACS_CyTOF_Clinical,  summary_reg_OLINK_FACS_CyTOF) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(Feature = gsub(Feature, pattern = "CD4mem_", replacement = "")) #%>%

p1 <- combined_regularizers %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "+", DataSet))) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  droplevels() %>%
  mutate(Feature = factor(Feature, levels = order_of_features)) %>%
  ggplot(aes(x = Feature, y = DataSet, fill = Occurrence)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("None" = "red", "LASSO only" = "#d9d9d9", "ELASTICNET only" = "#969696", "Both" = "#252525")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        strip.text.x = element_text(size = 8)) +
  labs(fill = "Method Occurrence")

p1
ggsave(p1, file = paste0("../Features_best_models_3models_reg.pdf"), width = 4, height = 2)
```

```{r}
cols_modality <- c("FACS" = "#c01c2c",
                   "OLINK" = "#2e3e9b",
                   "CyTOF" = "#5e2590",
                   "CyTOF a4b7" = "#ab43a8",
                   "Clinical" = "#656565")

combined_features <- rbind(feature_info_OLINK_FACS, feature_info_FACS_CyTOF_Clinical, feature_info_OLINK_FACS_CyTOF)

p1 <- combined_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  group_by(DataSet) %>%
  mutate(z_score_group = scale(absCoef)) %>%
  droplevels() %>%
  arrange(Origin, Feature) %>%
  mutate(Feature = factor(Feature, levels = c(order_of_features))) %>%
  ggplot(aes(x = Feature, y = z_score_group)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_boxplot(width = 0.5, outlier.size = 0.4) +
  ylab("Feature Importance") +
  xlab("Features") +
  scale_fill_manual(values = cols_modality) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8))
p1

p2 <- combined_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "+", DataSet))) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  mutate(Feature = gsub(Feature, pattern = "CD4mem_", replacement = "")) %>%
  dplyr::count(Feature, DataSet, Origin) %>%
  mutate(n_norm = n / 10) %>%
  arrange(Origin) %>%
  droplevels() %>%
  mutate(Feature = factor(Feature, levels = order_of_features)) %>%
  ggplot(aes(x = Feature, y = DataSet)) +
  geom_point(size = 3) +
  scale_color_manual(values = cols_modality) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8))

p1 <- p1 +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none")

p3 <- cowplot::plot_grid(p1, p2, labels = c('', ''), rows = 2, rel_heights = c(0.9, 0.8), align = 'v')
p3
ggsave(p3, file = paste0("../Features_best_models_3models.pdf"), width = 4, height = 2.5)
```

```{r}
sessionInfo()
```