### 2024-10-14 machine learning/ML_Treatment_efficacy_visualization.rmd

---
title: "R Notebook"
output: html_notebook
---

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpubr)

theme_figures <- function() {
  theme_minimal() %+replace%    #replace elements we want to change
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 8),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8),
          strip.text.x = element_text(size = 8)) }
```
# Model performances

## K-fold cross validation approach of

```{r auc_read_in}

# Data read in
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_treatment_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*pred_test.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]), row.names = 1))
  colnames(data) <- c('TrainingMeanAUC', 'TrainingStdAUC', 'TestingMeanAUC', 'TestingStdAUC')
  data <- t(round(data, 2))
  mat <- matrix(data, nrow = 2, ncol = 2)
  rownames(mat) <- c("MeanAUC", "SEMAUC")
  colnames(mat) <- c("Training", "Test")

  combined_data[[i]] <- mat
}
# Data tweaking
names(combined_data) <- gsub(pattern = "R_treatment_|_pred_test.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_accuracy <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "Set", "MeanAUC", 'SEMAUC')) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|LinearLASSO_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LASSO", Condition), "LinearLASSO",
                                       ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                              ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                                     ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                            ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinearLASSO", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none"))))))

```

```{r}
selected_models <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(DataSet %in% good_models) %>%
  filter(FeatureCutoff %in% c("none", "90%")) %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  arrange(SelectionCriteria) %>%
  slice_tail(n = 4) %>%
  pull(DataSet)
```

```{r}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_treatment_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*_coefficients.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_treatment_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LASSO", Condition), "LinearLASSO",
                                       ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                              ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                                     ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                            ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinearLASSO", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
                                                     mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))

# Code Figure 4E right panel
svg("../PredictiveModels/results/F_4E_right.svg", height = 1.55, width = 2)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% selected_models) %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>% filter(FinalDown == "none") %>%  group_by(Feature, DataSet) %>%
  mutate(Feature = factor(Feature, levels=c("Eosinophils","NK cells", "cMonocytes","Plasmablasts","cDCs", "DNER","CX3CL1", "CCL11","SCF", "CCL28","TNFB","ADA","TRANCE", "PD1", "CD103"))) %>%
  mutate(DataSet = factor(DataSet, levels = c("OLINK", "OLINK\nFACS", "FACS\nCyTOFab\nOLINK", "all\nCyTOFab"))) %>%
  dplyr::count(Feature, DataSet) %>%
  mutate(n_norm = n/10 ) %>%
  ggplot(aes(x= Feature, y = DataSet)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 6),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text.x=element_text(size = 6),
        strip.text.x = element_text(size = 8))
dev.off()

# Code Figure 4E left panel
svg("../PredictiveModels/results/F_4E_left.svg", height = 1.55, width = 2)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% selected_models) %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
  filter(FinalDown == "none") %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet))%>%
  group_by(DataSet) %>%
  mutate(z_score_group = scale(absCoef)) %>%
  ggplot(aes(x = reorder(Feature, -z_score_group, FUN = median), y = z_score_group)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_boxplot(width=0.5) +
  ylab("Z Score") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 6),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text.x=element_text(size = 6),
        strip.text.x = element_text(size = 8))

  dev.off()
```

`